---
title: "Assigment_2-knn"
output: html_document
date: "2026-10-05"
Author: "Vincent Fonte"
---

#Question 1 SPlitting the data 60/40
```{r}
library(caret)
library(class)
UniversalBank<-read.csv("UniversalBank.csv")
```
```{r}

UniversalBank$Education_1<-ifelse(UniversalBank$Education ==1,1,0)
 UniversalBank$Education_2<-ifelse(UniversalBank$Education ==2,1,0)
 UniversalBank$Education_3<-ifelse(UniversalBank$Education ==3,1,0)
 
bank_model<-subset(UniversalBank, select =-c(ID, ZIP.Code, Education))
X<- subset(bank_model, select = -Personal.Loan)
y <- bank_model$Personal.Loan

table(y)

```

```{r}
set.seed(43)
train_index<- sample(1:nrow(X),size = .60 * nrow(X))
 X_train<-X[train_index, ]
 y_train<-y[train_index]
  y_valid<-y[-train_index]
X_valid<- X[-train_index, ]
nrow(X_train)
nrow(X_valid)
length(y_train)
length(y_valid)
```

```{r}
preprocess <- preProcess(X_train, method =c("center", "scale"))
X_train_scaled<- predict(preprocess, X_train)
 X_valid_scaled<-predict(preprocess, X_valid)
```
```{r}
set.seed(43)
knn_1<-knn(train =X_train_scaled, test = X_valid_scaled, cl = y_train, k=1,prob=TRUE)
confusion_matrix_1<-table(
  predicted=knn_1,
  Actual=y_valid
  )
confusion_matrix_1
accuracy_1<- mean(knn_1 == y_valid)
accuracy_1
```
#Question 1 this custmer would be valdited as k=0 meaning the customer rejects the loan


# question 2 comparing different values of k=3,5,7,9

```{r}
 knn_3<- knn(train = X_train_scaled, test = X_valid_scaled,cl = y_train, k=3)

knn_5<- knn(train = X_train_scaled, test = X_valid_scaled,cl = y_train, k=5)
knn_7<- knn(train = X_train_scaled, test = X_valid_scaled,cl = y_train, k=7)

knn_9<- knn(train = X_train_scaled, test = X_valid_scaled,cl = y_train, k=9)
accuracy_3<-mean(knn_3==y_valid)
accuracy_5<-mean(knn_5==y_valid)
accuracy_7<-mean(knn_7==y_valid)
accuracy_9<-mean(knn_9==y_valid)

accuracy_3
accuracy_5
accuracy_7
accuracy_9


```

#Question 2:anwser Value of K 3 has the highest accuracy of .9630 this allows to not be overfitting or completely ignore data that can help us to predict

```{r}
confusion_matrix3<-table(predi=knn_3,Actual = y_valid)
confusion_matrix3
```
#anwser for 3 see chart above

```{r}
new_customer<- data.frame(Age=40, Experience=10, Income=84, Family=2, CCAvg=2, Education_1=0, Education_2=1, Education_3=0, Mortgage=0, Securities.Account=0, CD.Account=0, Online=1, CreditCard=1)

new_customer<-predict(preprocess, new_customer)
prediction <- knn(train = X_train_scaled, test = new_customer,cl = y_train, k=3)
prediction
```
#question 4 customer code is 0 this tells us that the customers will not accept the loan
```{r}
set.seed(43)
train_index_50<- sample(1:nrow(X), size= .50*nrow(X))
remaining_index<-setdiff(1:nrow(X), train_index_50)
valid_index30<- sample(remaining_index, size = .60 * length(remaining_index))
test_index20<-setdiff(remaining_index, valid_index30)

X_train_50<-X[train_index_50, ]
 y_train_50<-y[train_index_50]
 
 X_valid_30<- X[valid_index30, ]
 y_valid_30<- y[valid_index30]
 X_test_20<- X[test_index20, ]
 y_test_20<- y[test_index20]
 
 nrow(X_train_50)
 nrow(X_valid_30)
 nrow(X_test_20)
```
#question5 anwser

```{r}
preprocess_50<- preProcess(X_train_50, method =c("center", "scale"))
 X_train_50_scaled<- predict(preprocess_50, X_train_50)
X_valid_30_scaled<-predict(preprocess_50, X_valid_30)
X_test_20_scaled<-predict(preprocess_50, X_test_20)


```

```{r}
knn_train_50<- knn(train = X_train_50_scaled, test = X_train_50_scaled,cl = y_train_50, k=3)
confusion_train50<-table(predicted=knn_train_50,Actual=y_train_50)
confusion_train50

```
```{r}
accuracy_train50<-mean(knn_train_50==y_train_50)
accuracy_train50
```


```{r}
knn_valid_30<-knn(train=X_train_50_scaled,test=X_valid_30_scaled,cl=y_train_50,k=3)
confusion_val_30<-table(predicted=knn_valid_30,Actual=y_valid_30)
confusion_val_30
accuracy_val_30<-mean(knn_valid_30==y_valid_30)
accuracy_val_30
```
```{r}
knn_test_20<-knn(train=X_train_50_scaled,test=X_test_20_scaled,cl=y_train_50,k=3)
confusion_test_20<-table(predicted=knn_test_20,Actual=y_test_20)
confusion_test_20
accuracy_test_20<-mean(knn_test_20==y_test_20)
accuracy_test_20
```
#5 anwsers

